---
title: "linear models"
author: "Zakari L. Billo"
date: "2025-11-06"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.height = 7,
  fig.align = "center",
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

data + cleaning 

```{r}
nyc_airbnb = 
  nyc_airbnb |>
  mutate(
    stars = review_scores_location / 2 
  ) |>
  rename(
    borough = neighbourhood_group
  ) |>
  filter(borough != "Staten Island") |>
  select(price, stars, borough, room_type, neighbourhood)
```

do regression

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

do some additional cleaning  then refit

```{r}
nyc_airbnb = 
  nyc_airbnb |>
    mutate(
      borough = fct_infreq(borough),
      room_type = fct_infreq(room_type)
    )

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

summary of `lm`

```{r}
summary(fit)
names(summary(fit))
summary(fit)[["coefficients"]]
summary(fit)[["df"]]

fitted.values(fit)
resid(fit)
```

broom

```{r}
fit |>
  broom::tidy() |>
  mutate(
    term = str_replace(term, "borough", "Borough: ")) |>
  select(term, estimate, p.value) |>
  knitr::kable(digits = 3)
    
fit |>
  broom::glance()
```

## diagnostics

```{r}
nyc_airbnb |>
  modelr::add_residuals(fit) |>
  modelr::add_predictions(fit) |>
  filter(resid < 1000) |>
  ggplot(aes(x = resid)) +
  geom_histogram()

nyc_airbnb |>
  modelr::add_residuals(fit) |>
  modelr::add_predictions(fit) |>
  filter(resid < 1000) |>
  ggplot(aes(x = stars, y = borough)) +
  geom_violin()

nyc_airbnb |>
  modelr::add_residuals(fit) |>
  modelr::add_predictions(fit) |>
  filter(resid < 1000) |>
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```

what about adding categorical var (omnibus F test)

```{r}
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)

anova(fit_null, fit_alt) |>
  broom::tidy()
```

## interactions vs nested model

```{r}
fit_interactions = lm(price ~ stars*borough + room_type*borough, data = nyc_airbnb)

fit_interactions |>
  broom::tidy()
```

just look at BK first

```{r}
nyc_airbnb |>
  filter(borough == "Manhattan") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()

nyc_airbnb |>
  filter(borough == "Brooklyn") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()

nyc_airbnb |>
  filter(borough == "Queens") |>
  lm(price ~ stars + room_type, data = _) |>
  broom::tidy()
```

write a short function 

```{r}
lm_airbnb = function (df) {
  lm(price ~ stars + room_type, data = df)
}

nyc_airbnb |>
  filter(borough == "Queens") |>
  lm_airbnb() |>
  broom::tidy()
```

crease a list of dfs and iterate to fit the model each time 

```{r}
nested_lm_results = 
nyc_airbnb |>
  nest(data = -borough) |>
  mutate(
    fits = map(data, lm_airbnb),
    results = map(fits, broom::tidy)
  ) |> 
  select(borough, results) |>
  unnest(results) |>
  print()
```
do untidying for human readability

```{r}
nested_lm_results |>
  select(borough, term, estimate) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```

use an *anonymous* function instead of `lm_airbnb`

```{r}
anonymous_nested_lm_results = 
nyc_airbnb |>
  nest(data = -borough) |>
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)), # make internal, anonymous function w/ map statement
    results = map(fits, broom::tidy)
  ) |> 
  select(borough, results) |>
  unnest(results) |>
  print()
```

even more extreme

```{r}
manhattan_anal =
  nyc_airbnb |>
  filter(
    borough == "Manhattan"
  ) |>
  nest(data = -neighbourhood) |>
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(fits, broom::tidy)
  ) |>
  select(neighbourhood, results) |>
  unnest(results) |>
  print()
```

make plot

```{r}
manhattan_anal |>
  filter(term == "stars") |>
  mutate(
    neighbourhood = fct_reorder(neighbourhood, estimate)
  ) |> 
  ggplot(aes(x = neighbourhood, y = estimate)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

